---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `brfss2013`. Delete this note when before you submit 
your work. 

```{r load-data}
#load("brfss2013.RData")
```



* * *

## Part 1: Data

### Generalizability

The BRFSS study is a retrospective observationnal study.<br>
Random sampling is used to select adults in interviews (through random selction among a household or through random dialing), but only those who possess a landline phone or a cellular phone, and living in a private residence (household) or living in college housing. **So the results will be generalizable to all American adults living in a private residence or in college, and having a landline and/or a cellular phone.**<br>
In addition to this, modules about younger people (17yo or less) are optionnal, so not generalizable to the entire US population of 17yo and less.

### Causality

This is a study and not an experiment.<br>
So random assignment is not used, are people are asked about their personnal habits. **So the results will lead to correlation between variables but not to causality.**


* * *

## Part 2: Research questions

**Research quesion 1:**
- Is there a relationship between genre and general health level?
I find this question interesting, as it shows if genre (sex) as a correlation with overall feeling about one's health.

**Research quesion 2:**

**Research quesion 3:**


* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.

**Research quesion 1:**

First, I would like to create a new data frame which summarizes the sex var (genre of the person) and the genhlth var (general health), without any missing variable (NA). Let's call it health_genre.

```{r}
health_genre <- data_frame(sex = brfss2013$sex, health = brfss2013$genhlth)
health_genre <- na.omit(health_genre)
summary(health_genre)
```



**Research quesion 2:**

```{r}

```



**Research quesion 3:**

```{r}

```

